### 2014-09-09 paper2015asml.tex
initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
}

\author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small albert.author@papercept.net}}%
\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small b.d.researcher@ieee.org}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

This electronic document is a ÒliveÓ template. The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

This template provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

\section{PROCEDURE FOR PAPER SUBMISSION}

\subsection{Selecting a Template (Heading 2)}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
It may be used for A4 paper size if the paper size setting is suitably modified.

\subsection{Maintaining the Integrity of the Specifications}

The template is used to format your paper and style the text. All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them. You may note peculiarities. For example, the head margin in this template measures proportionately more than is customary. This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document. Please do not revise any of the current designations

\section{MATH}

Before you begin to format your paper~\cite{phillips2011sipp}, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.  Spell out units when they appear in text: Ò. . . a few henriesÓ, not Ò. . . a few HÓ.
\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use Ò(1)Ó, not ÒEq. (1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is . . .Ó

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ÒdataÓ is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
\item Do not confuse ÒimplyÓ and ÒinferÓ.
\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.

\end{itemize}

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ÒFig. 1Ó, even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

\section{CONCLUSIONS}

A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

The preferred spelling of the word ÒacknowledgmentÓ in America is without an ÒeÓ after the ÒgÓ. Avoid the stilted expression, ÒOne of us (R. B. G.) thanks . . .Ó  Instead, try ÒR. B. G. thanksÓ. Put sponsor acknowledgments in the unnumbered footnote on the first page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

%\begin{thebibliography}{99}
%
%\bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
%\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123Ð135.
%\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
%\bibitem{c4} B. Smith, ÒAn approach to graphs of linear forms (Unpublished work style),Ó unpublished.
%\bibitem{c5} E. H. Miller, ÒA note on reflector arrays (Periodical styleÑAccepted for publication),Ó IEEE Trans. Antennas Propagat., to be publised.
%\bibitem{c6} J. Wang, ÒFundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),Ó IEEE J. Quantum Electron., submitted for publication.
%\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
%\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ÒElectron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),Ó IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740Ð741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
%\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
%\bibitem{c10} J. U. Duncombe, ÒInfrared navigationÑPart I: An assessment of feasibility (Periodical style),Ó IEEE Trans. Electron Devices, vol. ED-11, pp. 34Ð39, Jan. 1959.
%\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, ÒA clustering technique for digital communications channel equalization using radial basis function networks,Ó IEEE Trans. Neural Networks, vol. 4, pp. 570Ð578, July 1993.
%\bibitem{c12} R. W. Lucky, ÒAutomatic equalization for digital communication,Ó Bell Syst. Tech. J., vol. 44, no. 4, pp. 547Ð588, Apr. 1965.
%\bibitem{c13} S. P. Bingulac, ÒOn the compatibility of adaptive controllers (Published Conference Proceedings style),Ó in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8Ð16.
%\bibitem{c14} G. R. Faulhaber, ÒDesign of service systems with priority reservation,Ó in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3Ð8.
%\bibitem{c15} W. D. Doyle, ÒMagnetization reversal in films with biaxial anisotropy,Ó in 1987 Proc. INTERMAG Conf., pp. 2.2-1Ð2.2-6.
%\bibitem{c16} G. W. Juette and L. E. Zeffanella, ÒRadio noise currents n short sections on bundle conductors (Presented Conference Paper style),Ó presented at the IEEE Summer power Meeting, Dallas, TX, June 22Ð27, 1990, Paper 90 SM 690-0 PWRS.
%\bibitem{c17} J. G. Kreifeldt, ÒAn analysis of surface-detected EMG as an amplitude-modulated noise,Ó presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
%\bibitem{c18} J. Williams, ÒNarrow-band analyzer (Thesis or Dissertation style),Ó Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
%\bibitem{c19} N. Kawasaki, ÒParametric study of thermal and chemical nonequilibrium nozzle flow,Ó M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
%\bibitem{c20} J. P. Wilkinson, ÒNonlinear resonant circuit devices (Patent style),Ó U.S. Patent 3 624 12, July 16, 1990. 
%
%
%
%
%
%
%\end{thebibliography}

\bibliography{IEEEabrv,max,arjun}
\bibliographystyle{IEEEtran}

\end{document}

### 2014-09-09 paper2015asml.tex
added some authors

@@
 \title{\LARGE \bf
-Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
+WORKING TITLE: Incorporating User Feedback in Motion Planning for Manipulators
 }
 
-\author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
-\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-        University of Twente, 7500 AE Enschede, The Netherlands
-        {\tt\small albert.author@papercept.net}}%
-\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
-        Dayton, OH 45435, USA
-        {\tt\small b.d.researcher@ieee.org}}%
+\author{Arjun Menon$^{1}$, Sachin Chitta$^{1}$, Mark Moll$^{2}$, Lydia Kavraki$^{2}$% <-this % stops a space
+\thanks{*This work was supported by ...}% <-this % stops a space
+\thanks{$^{1}$SRI International, Menlo Park, CA
+        {\tt\small arjun.menon@sri.com, sachin.chitta@sri.com}}%
+\thanks{$^{2}$Rice University, Department of Computer Science, Houston, TX
+        {\tt\small mmoll@rice.edu, kavraki@rice.edu}}%
 }

### 2014-10-02 paper2015asml.tex
more polish

@@
 %\usepackage{amssymb}  % assumes amsmath package installed
+\usepackage{wrapfig}
+\usepackage{graphicx,color}
+
+\usepackage{textcomp}
+\usepackage{amsmath}
+\usepackage{cases}
+\usepackage{xspace}
+%\usepackage{algorithm}
+%\usepackage{algorithmic}
+\usepackage{array}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+
+%EDITING COMMANDS
+\newcommand{\hilight}[1]{\colorbox{red}{#1}}
+\newcommand{\todo}{\colorbox{red}{TODO}}
+\newcommand{\tj}[1]{\ensuremath{\xi_\text{#1}}}
+
+\newcommand{\gta}[3]{pictures/example_traj/Good/#1-snapshot-#3-#2}
+\newcommand{\bta}[3]{pictures/example_traj/Bad/#1-snapshot-#3-#2}
+\newcommand{\fta}[3]{pictures/example_features/#1\_#2\_#3}
 
 \title{\LARGE \bf
-WORKING TITLE: Incorporating User Feedback in Motion Planning for Manipulators
+Towards a Data-driven Approach to Motion Planning
 }
 
-\author{Arjun Menon$^{1}$, Sachin Chitta$^{1}$, Mark Moll$^{2}$, Lydia Kavraki$^{2}$% <-this % stops a space
+\author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{1}$% <-this % stops a space
 \thanks{*This work was supported by ...}% <-this % stops a space
@@
         {\tt\small arjun.menon@sri.com, sachin.chitta@sri.com}}%
-\thanks{$^{2}$Rice University, Department of Computer Science, Houston, TX
-        {\tt\small mmoll@rice.edu, kavraki@rice.edu}}%
+\thanks{$^{2}$Independent Consultant,
+        {\tt\small }}%
 }
@@
 
-This electronic document is a ÒliveÓ template.
-The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.
-
+Co-robots, i.e. robots that work close to people, will need to account for the preferences and expectations of their human 
+co-workers in executing trajectories or actions. Consistent, legible and predictable trajectories are a key factor in making humans comfortable 
+around robots. In this work, we take a data-driven approach towards designing robot trajectories that are more acceptable 
+to human co-workers and observers. We use an online survey to ask people to rate multiple robot trajectories generated in a variety of environments. 
+We compute a large set of features for each trajectory, also taking into account environment information. We use a combination 
+of the features and the survey ratings to learn a classifier that predicts the rating for a new trajectory based on the learned 
+human-observer preferences. The classifier also helps identify and highlight the most important features 
+that influence people's ratings of the trajectories. Finally, we discuss how a data-driven approach using the 
+results of this analysis can be used to help design better trajectories that are more acceptable to people.
 \end{abstract}
@@
 
-This template provides authors with most of the formatting specifications needed for preparing electronic versions of their papers.
-All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings.
-Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example.
-Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided.
-The formatter will need to create these components, incorporating the applicable criteria that follow.
+Robotic motion generation in industrial environments does not take 
+into account any notion of the acceptability of trajectories for human observers. The need for more flexibility in manufacturing and logistics, 
+though, has resulted in a greater need for robots that can work beside humans. As robots and people start becoming co-workers, taking 
+human perception of robotic paths into account will become more important. Humans are able to predict and adjust their trajectories to 
+account for the expectations and trajectories of their co-workers easily. For human co-workers to be comfortable, robots will similarly 
+have to tailor their trajectories for the presence and actions of the people around them.
 
-\section{PROCEDURE FOR PAPER SUBMISSION}
+In this work, we take a first step towards the ultimate goal of generating robotic arm trajectories in cluttered environments that are acceptable for human observers and co-workers.
+We take a data-driven approach to this problem, using an online survey where human observers rate robot motion quality to generate a large dataset of human-observer preferences for robot trajectories.
+We use a feature set computed for each trajectory to then learn a relationship between the features in the trajectory and the rating that it gets from the human observers.
+A classifier learned in this manner serves two purposes: (a) it helps in identifying a set of features that are the most important in predicting human preferences for robot trajectories and (b) it could form the basis of a new approach to generating trajectories that is driven by learned human-observer preferences.
 
-\subsection{Selecting a Template (Heading 2)}
+This paper is organized as follows.
+In Section~\ref{sec:background}, we first describe related work in this area.
+In Section~\ref{sec:survey}, we present details of the online survey including information about the types of trajectories and environments used.
+Section~\ref{sec:raw_results} presents the results from the survey and analysis, including the trajectory features used for classification and the classification approaches used with the data.
+In Section~\ref{sec:conclusion}, we conclude with a discussion of how our results could be used to design motion planners that generate more human-acceptable trajectories.
+\begin{figure}
+\centering
+\begin{minipage}[][][s]{0.45\columnwidth}
+\centering
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{2}}
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{4}}
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}}
+(a)
+\end{minipage}
+\hfill
+\begin{minipage}[][][s]{0.45\columnwidth}
+\centering
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{2}}
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{4}}
+\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}}
+(b)
+\end{minipage}
+\caption{Example trajectory snapshots for motions that survey participants rated (a) highly and (b) poorly.
+In (b) the hand of the robot ducks under the table in between the first and middle snapshot.}\label{fig:example_trajectory}
+\end{figure}
 
-First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
-It may be used for A4 paper size if the paper size setting is suitably modified.
+\section{Background}
+\label{sec:background}
 
-\subsection{Maintaining the Integrity of the Specifications}
+\begin{figure}[b]
+\centering
+  \includegraphics[trim = 70mm 0mm 70mm 0mm, width=0.62\columnwidth, clip=true]{pictures/countertop_scene2}
+  \begin{minipage}[b][][s]{0.352\columnwidth}
+  \centering
+    \includegraphics[trim = 70mm 0mm 70mm 0mm, width=\textwidth, clip=true]{pictures/drawer_scene2}
+    \vfill
+    \includegraphics[trim = 30mm 0mm 30mm 0mm, width=\textwidth, clip=true]{pictures/industrial_scene2}
+  \end{minipage}
+  \hfill
+\caption{The scenes used in generating the robot motions to collect feedback from human observers.
+\emph{Drawer} (top right), \emph{Countertop} (left), and \emph{Industrial} (bottom right)}\label{fig:scenes}
+\end{figure}
 
-The template is used to format your paper and style the text.
-All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them.
-You may note peculiarities.
-For example, the head margin in this template measures proportionately more than is customary.
-This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations
+Robot motion has been primarily influenced by the need to generate smooth jerk-free trajectories for industrial robots that minimize wear and tear of internal parts.
+In recent years, there has been more interest in the generation of motions for robots working in human environments, i.e.
+environments with clutter and people working next to the robots.
+Most attention, though, has been focused on generating {\em collision-free} motions.
+Examples include the use of randomized planners~\cite{kuffner2000rrt}, trajectory optimization techniques~\cite{Mrinal:2011}, and deterministic search-based techniques~\cite{cohen2010search}.
 
-\section{MATH}
+Human motions have been the subject of recent study with the intent of determining whether there are certain motor or locomotor invariants that best describe human manipulation or locomotion.
+Human motion has been observed to be stereotypical~\cite{Atkeson:85}, i.e.
+learned motion patterns that are repeatedly used.
+This requires less planning of individual motions and contributes to making human motions more deterministic and predictable.
+The minimum jerk principle for human motions specifies that the trajectory of the human can best be represented as a motion that minimizes jerk of the hand~\cite{Flash:85}.
+Uno et.
+al.~\cite{Uno:89} used minimum torque change as a measure to explain the observed human motions.
+Minimizing jerk at the joint level~\cite{Rosenbaum:1995} has been another criterion used for human motion.
+Human motion has also been found to be optimized~\cite{Arechavaleta:2006}.
+Albrech et.
+al.~\cite{Albrecht:2011} studied the reaching motions of several subjects in table-setting situations and attempted to codify the motion using cost functions representing minimum-jerk hand motions, minimum-jerk joint motions, minimum torque change and a combination of all three.
+They found that no single combination of the cost functions explained the observed data completely.
 
-Before you begin to format your paper~\cite{phillips2011sipp}, first write and save the content as a separate text file.
-Keep your text and graphic files separate until after the text has been formatted and styled.
-Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph.
-Do not add any kind of pagination anywhere in the paper.
-Do not number text heads-the template will do that for you.
+Recent work has also addressed the task of making motions predictable (similar motions in similar environments and tasks) and/or legible (make it easier for human observers to understand the intent of the robot)~\cite{Beetz:2010}.
+Legibility makes it possible for human observers to understand clearly what a robot is intending to do, allowing them to modify their actions and motions if necessary.
+Predictability implies that the robot will always be consistent in its actions, removing the element of surprise for human observers.
+These concepts have been well-studies recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
+In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers.
 
-Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:
+There has been relatively little examination of how human observers perceive robot arm trajectories in environments with obstacles.
+There is no data available for how human observers react to such robot motions.
+It is not clear whether minimum-jerk trajectories (or other cost functions) are the best descriptors of motions in such environments or if there are other features that better explain human-observer preferences in such sitations.
+Our goal in this work is to collect the data needed to answer the query, "Which robot trajectories in cluttered environments do human-observers prefer".
+A second question that we would like to answer is "Why are certain robot trajectories preferred by human-observers?".
+This is a larger question that we aim to address in future work and instead we focus on,"Can we predict a human-observer preference for a given trajectory".
+The answer to this latter question, we hope, will eventually point us towards trajectories that are better accepted by human observers.
 
-\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract.
-Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined.
-Do not use abbreviations in the title or heads unless they are unavoidable.
+\section{Approach}
+\label{sec:survey}
+Our approach starts with an online survey designed to get information from multiple users about how well they like robot trajectories.
+In this section, we will describe in detail the infrastructure used to create the trajectories and movies that were used in the survey.
+We will also describe the online survey procedure in detail.
 
-\subsection{Units}
+\subsection{Planning System Infrastructure}
 
-\begin{itemize}
+\begin{wrapfigure}[13]{l}{.5\linewidth}
+\centering
+  \includegraphics[trim = 0mm 0mm 40mm 0mm, width=0.5\columnwidth, clip=true]{pictures/pr2_robot_4_downscale}
+\caption{{\footnotesize PR2 robot.}}
+\label{fig:PR2}
+\vspace{0.1in}
+\end{wrapfigure}
 
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.
-Spell out units when they appear in text: Ò.
-.
-.
-a few henriesÓ, not Ò.
-.
-.
-a few HÓ.
-\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)
+We extended the MoveIt!
+benchmarking architecture~\cite{cohen2012generic} with video recording capabilities to quickly generate a large number of trajectories.
+The MoveIt!
+architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries).
+It also allows for the use of any supported planning algorithms (planners).
+A combination of different scenes, motion queries, and planners was used to generate robot trajectories for evalutation.
+We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen (Countertop), an industrial workbench (Industrial) and a drawer on a table (Drawer).
+Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot (Fig.~\ref{fig:PR2}.
+The robot was positioned in the scenes manually and motion planning queries were constructed by hand.
+The positioning in the scenes was done with the expectation that the planning queries would not be impossible or trivial for the planning algorithms to solve.
 
-\end{itemize}
-
-\subsection{Equations}
-
-The equations are an exception to the prescribed specifications of this template.
-You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font).
-To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled.
-Number equations consecutively.
-Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop.
-To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents.
-Italicize Roman symbols for quantities and variables, but not Greek symbols.
-Use a long dash rather than a hyphen for a minus sign.
-Punctuate equations with commas or periods when they are part of a sentence, as in
-
-$$
-\alpha + \beta = \chi \eqno{(1)}
-$$
-
-Note that the equation is centered using a center tab stop.
-Be sure that the symbols in your equation have been defined before or immediately following the equation.
-Use Ò(1)Ó, not ÒEq.
-(1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is .
-.
-.Ó
-
-\subsection{Some Common Mistakes}
-\begin{itemize}
-
-\item The word ÒdataÓ is plural, not singular.
-\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
-\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
-\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
-\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
-\item Do not confuse ÒimplyÓ and ÒinferÓ.
-\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
-\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.
-
-\end{itemize}
-
-\section{USING THE TEMPLATE}
-
-Use this sample document as your LaTeX source file to create your document.
-Save this file as {\bf root.tex}.
-You have to make sure to use the cls file that came with this distribution.
-If you use a different style file, you cannot expect to get required margins.
-Note also that when you are creating your out PDF file, the source file is only part of the equation.
-{\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size.
-Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output.
-}
-
-It is impossible to account for all possible situation, one would encounter using \TeX.
-If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.
-
-\subsection{Headings, etc}
-
-Text heads organize the topics on a relational, hierarchical basis.
-For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic.
-If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced.
-Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.
-
-\subsection{Figures and Tables}
-
-Positioning Figures and Tables: Place figures and tables at the top and bottom of columns.
-Avoid placing them in the middle of columns.
-Large figures and tables may span across both columns.
-Figure captions should be below the figures; table heads should appear above the tables.
-Insert figures and tables after they are cited in the text.
-Use the abbreviation ÒFig.
-1Ó, even at the beginning of a sentence.
-
-\begin{table}[h]
-\caption{An Example of a Table}
-\label{table_example}
-\begin{center}
-\begin{tabular}{|c||c|}
+\begin{table*}[t]
+\centering
+\scalebox{0.9}{
+\begin{tabular}{|c||c|c|c|c|c|c|c|}
 \hline
-One & Two\\
+Efficiency  & Highly inefficient& Moderately inefficient& Slightly inefficient& Neutral & Slightly efficient& Moderately efficient& Highly efficient\\
 \hline
-Three & Four\\
+Elegance    & Highly awkward& Moderately awkward& Slightly awkward& Neutral & Slightly elegant& Moderately elegant& Highly elegant\\
+\hline
+Smoothness  & Highly rough& Moderately rough& Slightly rough& Neutral & Slightly smooth& Moderately smooth& Highly smooth\\
+\hline
+Overall     & Extremely poor& Moderately poor& Somewhat poor& Neutral & Somewhat good& Moderately good& Extremely good\\
 \hline
 \end{tabular}
-\end{center}
+}
+\caption{The rating categories and the associated scale presented to the survey respondent}
+\label{tab:rating_scales}
+\end{table*}
+
+%\begin{figure}[b]
+%    \includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/designing_queries_moveit}
+%    \caption{Screenshot of the planning query design process which involved hand-picking start and goal query poses for the various scenes manually to ensure they were neither trivial nor impossible for the planning algorithms to solve.
+This was accomplished through use of RVIZ and interactive markers.}
+%    \label{fig:design}
+%\end{figure}
+
+The planning algorithms that we used are RRT-Connect~\cite{kuffner2000rrt} and RRT*.
+RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: PathLength and MaxMinClearance.The objective for PathLength minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments).
+The objective for MaxMinClearance computes uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize.
+Both planners are implemented in the OMPL library~\cite{OMPL:2012}.
+The planned trajectories are post-processed by an iterative-time parameterization method that time-parameterizes the trajectories while taking into account velocity and accelerations.
+The infrastructure we designed around MoveIt!
+allows us to automatically generate movies of the generated trajectories using the ROS Visualizer (Rviz).
+We generated movies from different viewpoints to maximize the visibility of the motion to human observers reviewing it.
+A total of 103 different trajectories were generated in this manner.
+
+\subsection{Online Survey}
+
+We conducted an online survey\footnote{The online survey was conducted after gaining a waiver from SRI's IRB after a review of our survey procedure.
+Our survey followed standard guidelines for privacy of respondent information when conducting a survey on Mechanical Turk.} where we asked human observers to rate a variety of trajectories of the PR2 robot.
+The survey was built on Amazon's Mechanical Turk~\cite{paolacci2010running}, which has become a useful tool for large online data gathering and crowd-sourced work.
+Each respondent in the survey was presented with a set of demographic questions, a video, and four scales to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
+Each respondent was expected to rate multiple trajectories but could not rate a trajectory more than once.
+Each video has four different views of the robot displayed simultaneously.
+Respondents were asked demographic questions about familiarity with video games, familiarity with robotics, age group and highest education level.
+The demographic information was not used to filter the responses we collected.
+
+Mechanical Turk allows monetary compensation on completion of the survey, which we set to 15\textcent~per survey.
+We estimated the time that each respondent would take to rate a single trajectory as 1-2 minutes.
+Additionally, the respondents were allowed to replay the trajectory video repeatedly.
+The maximum possible payout for each respondent is thus 15.45\$ (if each survey took 1.5 minutes to complete then the wage is approximately 6\$ per hour).
+This is standard compensation for quality work on Mechanical Turk for tasks of this type, as was pointed out to us by several veteran Mechanical Turk surveyors.
+
+Respondents on Mechanical Turk are only paid if the survey organizer approves their response.
+A control question, asking the respondent to identify the known action of a robot in a video, served as the first satisficing trap to ensure that respondents were truthfully attempting the survey (see~\cite{krosnick1991response}~\cite{krosnick1996satisficing} for more information on satisficing in survey responses).
+Survey responses were also manually examined to assess whether respondents were responding honestly.
+We looked at whether respondent responses were outliers when compared to the rest of the group.
+We also examined the average time taken by each respondent to complete the survey.
+Overall, we found that all the respondents had made a honest attempt at filling out the rating and we did not reject any responses.
+
+%\begin{figure}
+%    \frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_survey_screenshot}}
+%    \caption{The main survey question which prompts the respondent for a rating in four motion characteristics (efficiency, elegance, smoothness and overall impression) after presenting a video that demonstrates the motion from four clearly visible vantage points for any sample trajectory.}
+%    \label{fig:survey_question}
+%\end{figure}
+
+%\begin{figure}
+%\frame{\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\columnwidth]{pictures/amazon_control_question}}
+%\caption{The survey control question which prompts the respondent about the action of the robot in the accompanying video.
+The question we adopted was to ask the particpant which colored cube the robot touched.}
+%\label{fig:control_question}
+%\end{figure}
+
+\section{Survey Results and Analysis}
+\label{sec:raw_results}
+The survey resulted in at least 20 ratings for each trajectory (so there were a total of 2060 ratings from participants in the survey).
+In this section, we will first analyze the raw results from the survey (which are interesting in themselves).
+Subsequently, we will present a set of features that were used to represent individual trajectories.
+These features will be used in the next section to learn the human-observer ratings for robot trajectories.
+\subsection{Raw Survey Results}
+\begin{figure}[h]
+  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
+  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
+  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
+  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
+  \caption{Histogram of averaged ratings for the 103 different trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Inelegant, Smooth vs. Rough, Overall Rating}
+\label{fig:survey_raw}
+\end{figure}
+The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of averaged responses for each of the four survey questions across the 103 trajectories that were rated by users.
+It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well.
+An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall).
+Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes.
+Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$).
+Confidence intervals for the overall rating of each trajectory are shown in Fig.~\ref{confidence}.
+The small spread in the confidence intervals clearly indicates that different users had similar opinions about the same trajectory.
+
+\begin{table}
+\centering
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+& Efficient & Elegant & Smooth & Overall \\ \hline
+Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
+Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
+Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
+Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
+\end{tabular}
+\caption{Correlation between the four attributes in the survey.}
+\label{tab:correlation}
 \end{table}
 
-   \begin{figure}[thpb]
+\subsection{Trajectory Features}
+\label{subsec:traj_feat}
+
+The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory.
+These make up our unary features computed only on the input trajectory.
+We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the "expected" path.
+We go over the multi-point features in the next section.
+First, we detail the features we computed on each trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$.
+Note that our features are generic, i.e.
+they can be computed in the same manner for any robot arm trajectory.
+We believe that our methods should be easily extensible to robots other than the PR2 as well and individual trajectory classifiers could be built for each different robot arm in a similar manner.
+
+{\bf Path length:}
+%The length of the trajectory \tj{} computed as the sum over trajectory segments of the euclidean distance between end points of the segment, or
+%\begin{equation}
+%\text{Length} = \sum_{i=0}^{n-1} \| q_\text{i}-q_\text{i+1} \|
+%\end{equation}
+We compute path length for the joint-space trajectory and for three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links of the right arm (the moving arm).
+We also compute the length of the trajectory in angular distance by summing over the waypoints $\arccos( \langle Q_\text{i},Q_\text{i+1}\rangle^2-1 )$ where $Q_\text{i}$ is the quaternion orientation for the $i$-th waypoint (referred to later as the "quat\_dist" feature).
+
+{\bf Clearance:}
+The clearance of a trajectory \tj{} is the average over the waypoints of the minimum distance of configurations to the closest obstacle.
+%\begin{equation}
+%\text{Clearance} = \frac{1}{n} \sum_{i=0}^n \min_{o \in O} \text{dist}(q_i, o)
+%\end{equation}
+%where $O$ represents the set of obstacles in the scene and $\text{dist}(q_i, o)$ calculates the closest distance between obstacle $o$ and configuration $q_i$.
+
+{\bf Smoothness:}
+The smoothness of a trajectory is computed by averaging the included angle of three consecutive waypoint configurations, for all triplets of waypoints in \tj{}.
+The angle is computed using joint-space euclidean distance between configurations as the length of sides of the triangle as described in~\cite{cohen2012generic}.
+%\begin{equation}
+%\begin{aligned}
+%&\text{Smoothness} = \frac{1}{n-2} \sum_{i=1}^{n-1} \mathellipsis\\
+%&\left[ \pi- \text{angle}(\|q_\text{i-1}-q_\text{i}\|,\|q_\text{i}-q_\text{i+1}\|,\|q_\text{i+1}-q_\text{i-1}\|) \right]\\
+%\end{aligned}
+%\end{equation}
+%where the included angle is computed
+%\begin{equation}
+%\text{angle}(a,b,c) = \arccos (a^2+c^2-b^2)/2ac
+%\end{equation}
+We compute smoothness for both the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.
+
+{\bf Maximum Acceleration:}
+We compute the absolute maximum acceleration for the joint-space trajectory \tj{} and for workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.
+%For a joint-space trajectory \tj{} consisting of waypoints, the maximum acceleration is computed by finding the accelerations at the knot points of splines fit to each joint's trajectory. 
+%\begin{equation}
+%\text{Max Acceleration} = \max_{i={0,\mathellipsis,n}} {\ddot{q}_i}^T\ddot{q}_i
+%\end{equation}
+
+{\bf Maximum Radius of Curvature: }
+Curvature at a given waypoint is computed by splines to the workspace trajectory.
+We compute maximum curvature for only the workspace trajectories $\xi_\text{elbow}$, $\xi_\text{wrist}$, and $\xi_\text{tip}$.
+% for each coordinate and computing from the higher order derivatives
+%\begin{equation}
+%R_i = \frac{\|\dot{q}_i\|^3}{\|\dot{q}_i \times \ddot{q}_i\|}
+%\end{equation}
+%where velocity $\dot{q}_i$ and acceleration $\ddot{q}_i$ are computed at the waypoint $i$ from differentiating the interpolating spline of the work-space trajectory.
+Alternatively, the higher order derviatives can be %computed using finite differencing methods.
+
+{\bf Joint Limit Distance: }
+The maximin distance to joint limits is the maximum over waypoints, of the miniminum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$.
+
+\subsection{Multi-Trajectory Features}
+\label{subsec:multi_traj_feat}
+We also wanted to design features that captured the difference between the planned trajectory and an expected path for the end-effector in the environment.
+The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D voxel approximation of the planning problem.
+The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty.
+Any voxel that is partially occupied is considered to be occupied.
+A 26-connected grid was used in the planning process to find a path in the voxel grid for a cubical shape moving from the start position of the end-effector of the arm to the desired goal position.
+Essentially, this would be the optimal path followed by the end-effector from start to goal if it were modeled as a free-standing cube.
+Next, we compute comparison features between $\xi_{\text{BFS}}$ and three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links.
+The comparison features computed between each pair of trajectories are the Hausdorff distances~\cite{dubuisson1994modified} and Dynamic Time Warping~\cite{senin2008dynamic} distances.
+
+%The start and goal voxel is simply the $(x,y,z)$ of the end-effector tool-frame. The free voxels then make up our vertices $V_\text{grid}$ for our graph.
+
+%The edges $E_\text{grid}$ used in this search are the 26-connected grid of adjacent neighboring voxels.
+For example, some valid adjacent voxels can be computed by adding $(\Delta x,\Delta y,\Delta z)$ to the current %coordinates, where $\Delta x= \Delta y= \Delta z= \{\pm1,0\}$.
+An edge only exists in this graph if the ending voxel is collision.
+
+%The order the neighboring voxels are generated by the breadth-first search is done such that neighboring voxels that share \emph{faces} are pushed onto the queue before those that share an \emph{edge}, which are then %added before those that share a \emph{corner} with the generating state.
+This lets us use a simple BFS to search the graph with unit cost edges.
+
+%Using the graph $G=(V_\text{grid}, E_\text{grid})$, we perform breadth-first search to get a collision free path from $V_\text{start}$ to $V_\text{end}$.
+The resulting path is given a nominal timing and we refer to it as %$\xi_{\text{BFS}}$.
+
+{\bf Hausdorff Distance:}
+The Hausdorff distance $H_\text{AB}$ between trajectory $\xi_\text{A}$ and trajectory $\xi_\text{B}$ can be computed as follows:
+\begin{equation}
+\begin{aligned}
+H_\text{AB} &= \max (h_{\text{AB}}, h_{\text{BA}}) \\
+h_{\text{AB}} &= \max_{q_\text{A} \in \xi_\text{A}} \min_{q_\text{B} \in \xi_\text{B}} \text{d}(q_\text{A}, q_\text{B})\\
+h_{\text{BA}} &= \max_{q_\text{B} \in \xi_\text{B}} \min_{q_\text{A} \in \xi_\text{A}} \text{d}(q_\text{B}, q_\text{A})\\
+\end{aligned}
+\end{equation}
+
+where $q_\text{A}$ is a configuration (or in this case an end effector position) at the waypoint in trajectory \tj{A} and $\text{d}(q_\text{i}, q_\text{j})$ is the euclidean distance between the trajectory waypoints.
+
+We also use the Interpolation-based Modified Hausdorff distance~\cite{dubuisson1994modified}~\cite{chen2013dynamic}, which is defined as:
+
+\begin{equation}
+\text{imh}_\text{AB} = \frac{1}{n_A} \sum_{q_\text{A} \in \xi_\text{A}} \min_{\overline{q_\text{B,i}} \in \xi_\text{B}} \text{d}(q_\text{A}, \overline{q_\text{B,i}})
+\end{equation}
+
+where $n$ is the number of waypoints, $\overline{q_\text{B,i}}$ is the i-th segment in \tj{B}, and $\text{d}(q, \overline{q})$ is the shortest distance between the point and the line segment.
+We calculate this distance only one way, for $(\tj{elbow},\tj{BFS})$, $(\tj{wrist},\tj{BFS})$.
+and $(\tj{tip},\tj{BFS})$.
+
+{\bf Dynamic Time Warping:}
+
+Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}.
+It provides a way to compute a distance despite differences in timing of the trajectories as well.
+DTW computes a mapping of a subset of the waypoints in $\xi_\text{A}$ to a subset of the waypoints in $\xi_\text{B}$ which constitutes the warping path, or
+
+\begin{equation}
+\xi_\text{DTW} = \{ (0, 0), (a_i, b_j), \mathellipsis ,(n_A, n_B) \}
+\end{equation} where $n_A = |\tj{A}|$ and $n_B = |\tj{B}|$ which are the number of waypoints in the trajectory.
+More on the construction of the \tj{DTW} can be read in~\cite{senin2008dynamic}\cite{chen2013dynamic}.
+The DTW feature is the cost of the warping path.
+
+%The path is constrained to map the first waypoints and last waypoints together, which makes up a boundary condition.
+Finally, the warping path is continuous and monotone such that each coordinate of the waypoint in the warping path is between 0 and 1 of the coordinate of the preceding waypoint in the path, or
+%
+%\begin{equation}
+%\forall_{(a_i,b_i) \in \xi_\text{DTW}} 
+%  \begin{cases}
+%      0 \leq a_{i} - a_{i-1} \leq 1 \\
+%      0 \leq b_{i} - b_{i-1} \leq 1 \\
+%  \end{cases}
+%\end{equation}
+%
+%Computing this warping path cost is done by framing it as a dynamic programming problem where the cumulative DTW cost of a cell is done on a $n_B+1 \times n_B+1$ array and initialized such that
+%
+%\begin{equation}
+%\begin{aligned}
+%\text{DTW}[0,0] &= 0 \\
+%\forall_{i=\{1,\mathellipsis,n_A\}} \text{DTW}[i,0] &= \inf\\
+%\forall_{j=\{1,\mathellipsis,n_B\}} \text{DTW}[0,j] &= \inf\\
+%\end{aligned}
+%\end{equation}
+%
+%and the algorithm iterates over $i=\{1,\mathellipsis,|\tj{A}|\}$ and $j=\{1,\mathellipsis,|\tj{B}|\}$ and updates the the cell at an index $(i,j)$ with the rule
+%
+%\begin{equation}
+%\text{DTW}[i,j] = \text{d}_{i,j} + \min 
+%  \begin{cases}
+%    \text{DTW}[i-1,j] \\
+%    \text{DTW}[i,j-1] \\
+%    \text{DTW}[i-1,j-1] ) \\
+%  \end{cases}
+%\end{equation}
+%
+%where $\text{d}_{i,j}$ is the euclidean distance between waypoint $i$-th waypoint of \tj{A} and $j$-th waypoint of \tj{B}.
+After computing this array, $\text{DTW}[n_A, n_B]$ gives the cost or distance between \tj{A} and \tj{B}.
+
+%\begin{figure*}[t]
+%\begin{center}
+%\begin{tabular}{cccc}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_59}{3}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_59}{3}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_59}{3}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_59}{3}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_59}{3}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_22}{2}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_22}{2}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_22}{2}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_22}{2}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_22}{2}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_58}{3}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_58}{3}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_58}{3}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_58}{3}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_38_58}{3}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_36}{4}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_36}{4}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_36}{4}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_36}{4}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_36}{4}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_41}{2}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_41}{2}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_41}{2}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_41}{2}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_39_41}{2}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_15}{2}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_15}{2}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_15}{2}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_15}{2}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_15}{2}{5}}\\
+%
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_16}{3}{1}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_16}{3}{2}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_16}{3}{3}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_16}{3}{4}}
+%\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.19\textwidth]{\gta{2014_Sep_17_18_40_16}{3}{5}}\\
+%
+%bad trajs
+%{2014_Sep_17_18_38_44}{1}
+%{2014_Sep_17_18_39_17}{3}
+%{2014_Sep_17_18_40_06}{1}
+%{2014_Sep_17_18_40_21}{2}
+%
+%\end{tabular}
+%\end{center}
+%\caption{Snapshots of favorably rated trajectories on all four categories by a large proportion of survey users}
+%\label{fig:good_traj}
+%\end{figure*}
+
+\begin{table}
+\label{tab:feature_description}
+\scalebox{0.9}{
 \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Inductance of oscillation winding on amorphous
-       magnetic core versus DC bias magnetic field}
-      \label{figurelabel}
+\begin{tabular}{p{0.3\columnwidth}|p{0.2\columnwidth}|p{0.5\columnwidth}}
+\hline
+Feature Name & Metric & Feature Description \\
+\hline
+path\_length, smoothness, clearance, max\_lin\_acc, joint\_limit\_distance & Joint-space & Features discussed in Section~\ref{subsec:traj_feat} computed on joint-space trajectory \tj{}\\
+\hline
+[link]\_length, [link]\_smoothness, [link]\_max\_curvature, [link]\_lin\_acc, [link]\_quat\_dist & Cartesian & Features discussed in Section~\ref{subsec:traj_feat} computed on 3D cartesian path traced by [link] in \tj{}\\
+\hline
+h\_[link]\_bfs, h\_[link1]\_[link2], H\_[link]\_bfs, H\_[link1]\_[link2] & Cartesian & Assymetric ($h$) or symmetric ($H$) hausdorff distance between 3D cartesian path traced by [link] in \tj{} w.r.t BFS path or between 3D Cartesian path traced by [link1] and [link2] in \tj{}\\
+\hline
+[link]\_bfs\_dtw, [link1]\_[link2]\_dtw & Cartesian & DTW cost between path traced by [link] in \tj{} and the BFS path, or between 3D-cartesian-path traced by [link1] and [link2] in \tj{}\\
+\hline
+\end{tabular}
+}
+\caption{Decoding of feature names into descriptions}
+\end{table}
+
+\section{Classification}
+The survey data provides a great platform to answer the question, ``Can we predict a human-observer preference for a given trajectory?''.
+Our approach to this question is to let the data drive us towards the answer instead of trying to hand-design a cost function or set of features that offer the best explanation.
+Towards this end, we chose to learn a classifier based on the available data to help predict the human-observer rating for a new trajectory.
+The 103 data points (corresponding to the 103 distinct trajectories) were broken up into 3 classes based on average overall ratings:
+\begin{itemize}
+\item “low” - 11 data points with ratings $\in (1,3]$,
+\item “medium”  - 29 data points with ratings $\in (3,5]$,
+\item “high” – 63 data points with rating $\in (5,7]$.
+\end{itemize}
+The data was divided into two samples, a training sample and a test sample, with 1/3rd of each class (overall rating) used to construct the test data sample and the remaining 2/3rd data points within each class used to train the various models.
+This results in a training set consisting of 68 data points and test set consisting of 35 data points.
+Table~\ref{tab:test_training} shows how the average overall ratings are distributed across training and test sets.
+\begin{table}[h]
+\begin{centering}
+\begin{tabular}{|c|c|c|c|}
+\hline
+ & low (1,3] & medium (3,5] & high (5,7] \\ \hline
+Training & 7 & 19 & 42 \\ \hline
+Test & 4 & 10 & 21 \\ \hline
+\end{tabular}
+\caption{Distribution between training and test samples.}
+\label{tab:test_training}
+\end{centering}
+\end{table}
+
+In all, we utilized three different classification frameworks to predict overall ratings - decision trees, random forests and boosting.
+As a base framework we first construct a decision tree model.
+Decision trees for classification provide a good visual representation especially in the presence of complex features, where the relationships between predictors and response is not immediately obvious.
+The features of interest in this tree include the DTW distance computed for the wrist (vs.
+the tool path) denoted by wrist\_tool\_dtw, the Hausdorf distance computed for the end-effector and the max curvature computed for the elbow.
+Using 10-fold cross-validation, we also design a simpler tree that has just three terminal nodes with the wrist\_tool\_dtw feature serving as the key discriminant among the classes.
+This tree has an accuracy of 86\% on the test data.
+Large values of the wrist\_tool\_dtw feature correspond to extra rotations of the wrist as the tool is moving towards the goal - such trajectories are rated as low quality by the human observers.
+The second classification framework we use is a random forest ensemble of trees.
+This involves randomly dividing the data into several bootstrap training samples and building a tree on each bootstrap sample.
+Within each boostrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split.
+This brings additional randomness within each tree.
+The class is chosen by majority vote across all trees.
+For every tree constructed from a bootstrapped training sample there is one Out of Bag (OOB) error which can be thought of as a test error for every tree.
+It is estimated internally, during the run.
+On test data, the random forest method returns 83\% accuracy.
+
+Fig.~\ref{fig:feature_importance_1} and Fig.~\ref{fig:feature_importance_2} show two measures of variable (or feature) importance.
+The first measure shows the mean decrease in accuracy for each feature.
+The importance of a feature is determined by how much the accuracy of the random forest is affected by comparing, for each tree, the OOB error with the error obtained after permuting each feature.
+The differences are averaged across trees and normalized.
+The second measure is based on the Gini index and measures the total decrease in node heterogeneity/impurities from splitting on the feature.
+For both measures higher values denote that the features are more important.
+Again, the wrist\_tool\_dtw feature stands out as one of the more important features in building the classifer.
+Two path lengths tool\_length (the distance traveled by the tool) and path\_length (the total distance traveled in joint space) also serve as important features - in most environments the total distance traveled by the arm between start and goal locations is expected to be small.
+The tool\_quat\_dist and wrist\_quat\_dist measure the rotation of the end-effector and it seems natural that excessive rotations of the end-effector would contribute to trajectories getting a bad rating.
+It is also clear that the BFS path, which serves as the nominal path for computing both the DTW and Hausdorf distance, is a good approximation of the path expected by the human observer for the end-effector of the robot.
+
+The third classification framework that we use is Boosting.
+In the Boosting  framework, the trees are grown sequentially, using information (weights) from previously grown trees.
+The algorithm adaptively changes the distribution of the training data by iteratively reweighting each observation in each iteration.
+The weights are updated using (Breiman’s) weight updating coefficient which assigns higher weights to the wrongly classified observations and lower weights to the correctly classified examples.
+In every iteration, the  classifier thus focuses on the most difficult examples.
+Using 10-fold cross validation with all the available data, we obtain a classification accuracy of 76.7\%.
+Fig.~\ref{fig:gini_boosting} shows the relative importance of different features using the Boosting classifier.
+7 of the top 10 features on this list are also part of the top 10 features using random forests.
+
+\begin{figure}[t]
+  \centering{\includegraphics[width=0.99\linewidth]{pictures/meandecrease}}
+  \caption{Random Forest: Mean decrease in accuracy for each feature.}
+\label{fig:feature_importance_1}
 \end{figure}
 
-Figure Labels: Use 8 point Times New Roman for Figure labels.
-Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader.
-As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ.
-If including units in the label, present them within parentheses.
-Do not label axes only with units.
-In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ.
-Do not label axes with a ratio of quantities and units.
-For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó
+\begin{figure}[t]
+ \centering{\includegraphics[width=0.99\linewidth]{pictures/gini}}
+  \caption{Random Forest: Gini index (higher values indicate more important features).}
+\label{fig:feature_importance_2}
+\end{figure}
 
-\section{CONCLUSIONS}
+\begin{figure}[t]
+ \centering{\includegraphics[width=0.99\linewidth]{pictures/gini}}
+  \caption{Boosting: Gini index (higher values indicate more important features).}
+\label{fig:gini_boosting}
+\end{figure}
 
-A conclusion section is not required.
-Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion.
-A conclusion might elaborate on the importance of the work or suggest applications and extensions.
-
-\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                   % on the last page of the document manually. It shortens
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{APPENDIX}
-
-Appendixes should appear before the acknowledgment.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{ACKNOWLEDGMENT}
-
-The preferred spelling of the word ÒacknowledgmentÓ in America is without an ÒeÓ after the ÒgÓ.
-Avoid the stilted expression, ÒOne of us (R.
-B.
-G.) thanks .
-.
-.Ó  Instead, try ÒR.
-B.
-G.
-thanksÓ.
-Put sponsor acknowledgments in the unnumbered footnote on the first page.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.
-
-%\begin{thebibliography}{99}
-%
-%\bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
-%\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123Ð135.
-%\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-%\bibitem{c4} B. Smith, ÒAn approach to graphs of linear forms (Unpublished work style),Ó unpublished.
-%\bibitem{c5} E. H. Miller, ÒA note on reflector arrays (Periodical styleÑAccepted for publication),Ó IEEE Trans. Antennas Propagat., to be publised.
-%\bibitem{c6} J. Wang, ÒFundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),Ó IEEE J. Quantum Electron., submitted for publication.
-%\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-%\bibitem{c8} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ÒElectron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),Ó IEEE Transl.
-J.
-Magn.Jpn., vol.
-2, Aug.
-1987, pp.
-740Ð741 [Dig.
-9th Annu.
-Conf.
-Magnetics Japan, 1982, p.
-301].
-%\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-%\bibitem{c10} J. U. Duncombe, ÒInfrared navigationÑPart I: An assessment of feasibility (Periodical style),Ó IEEE Trans. Electron Devices, vol. ED-11, pp. 34Ð39, Jan. 1959.
-%\bibitem{c11} S.
-Chen, B.
-Mulgrew, and P.
-M.
-Grant, ÒA clustering technique for digital communications channel equalization using radial basis function networks,Ó IEEE Trans.
-Neural Networks, vol.
-4, pp.
-570Ð578, July 1993.
-%\bibitem{c12} R. W. Lucky, ÒAutomatic equalization for digital communication,Ó Bell Syst. Tech. J., vol. 44, no. 4, pp. 547Ð588, Apr. 1965.
-%\bibitem{c13} S.
-P.
-Bingulac, ÒOn the compatibility of adaptive controllers (Published Conference Proceedings style),Ó in Proc.
-4th Annu.
-Allerton Conf.
-Circuits and Systems Theory, New York, 1994, pp.
-8Ð16.
-%\bibitem{c14} G. R. Faulhaber, ÒDesign of service systems with priority reservation,Ó in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3Ð8.
-%\bibitem{c15} W. D. Doyle, ÒMagnetization reversal in films with biaxial anisotropy,Ó in 1987 Proc. INTERMAG Conf., pp. 2.2-1Ð2.2-6.
-%\bibitem{c16} G.
-W.
-Juette and L.
-E.
-Zeffanella, ÒRadio noise currents n short sections on bundle conductors (Presented Conference Paper style),Ó presented at the IEEE Summer power Meeting, Dallas, TX, June 22Ð27, 1990, Paper 90 SM 690-0 PWRS.
-%\bibitem{c17} J. G. Kreifeldt, ÒAn analysis of surface-detected EMG as an amplitude-modulated noise,Ó presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-%\bibitem{c18} J. Williams, ÒNarrow-band analyzer (Thesis or Dissertation style),Ó Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-%\bibitem{c19} N. Kawasaki, ÒParametric study of thermal and chemical nonequilibrium nozzle flow,Ó M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-%\bibitem{c20} J. P. Wilkinson, ÒNonlinear resonant circuit devices (Patent style),Ó U.S. Patent 3 624 12, July 16, 1990. 
-%
-%
-%
-%
-%
-%
-%\end{thebibliography}
-
+We would like to thank Leila Takayama for valuable advice on structuring the online survey. We would like to thank Lydia Kavraki and Mark Moll for valuable discussions and suggestions. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{IEEEabrv,max,arjun}

### 2014-10-02 paper2015asml.tex

@@
 \usepackage{graphicx,color}
+\usepackage{adjustbox}
 
@@
 \newcommand{\bta}[3]{pictures/example_traj/Bad/#1-snapshot-#3-#2}
-\newcommand{\fta}[3]{pictures/example_features/#1\_#2\_#3}
+\newcommand{\fta}[3]{pictures/example_features/#1_#2_#3}
+
+\newcommand*\rot{\rotatebox{90}}
 
@@
 
+\begin{figure}
+\centering
+\begin{tabular}{c|p{0.4\columnwidth}p{0.4\columnwidth}}
+& Low feature-value & High feature-value \\
+\hline
+
+\rot{wrist\_tool\_dtw} & 
+\adjincludegraphics[trim={.1\width} {.2\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_tool_dtw}{low}{0}} & 
+\adjincludegraphics[trim={.2\width} {.0\height} {.2\width} {.15\height}, clip, width=0.4\columnwidth]{\fta{wrist_tool_dtw}{high}{0}} \\
+
+\rot{tool\_length} & 
+\adjincludegraphics[trim={.1\width} {.2\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{tool_length}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.015\height}, clip, width=0.4\columnwidth]{\fta{tool_length}{high}{0}} \\
+
+\rot{tool\_quat\_dist} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{tool_quat_dist}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{tool_quat_dist}{high}{0}} \\
+
+\rot{wrist\_bfs\_dtw} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_bfs_dtw}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_bfs_dtw}{high}{0}} \\
+\end{tabular}
+
+%\begin{minipage}[][][s]{0.45\columnwidth}
+%\centering
+%\adjincludegraphics[trim={.1\width} {.2\height} {.25\width} {.0\height}, clip, width=\textwidth]{\fta{tool_length}{low}{0}}
+%\adjincludegraphics[trim={.1\width} {.0\height} {.25\width} {.0\height}, clip, width=\textwidth]{\fta{tool_quat_dist}{low}{0}}
+%\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=\textwidth]{\fta{wrist_bfs_dtw}{low}{0}}
+%%\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\textwidth, clip=true]{\fta{}{low}{0}}
+%(a)
+%\end{minipage}
+%\hfill
+%\begin{minipage}[][][s]{0.45\columnwidth}
+%\centering
+%\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=\textwidth]{\fta{tool_length}{high}{0}}
+%\adjincludegraphics[trim={.1\width} {.0\height} {.25\width} {.0\height}, clip, width=\textwidth]{\fta{tool_quat_dist}{high}{0}}
+%\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=\textwidth]{\fta{wrist_bfs_dtw}{high}{0}}
+%%\includegraphics[trim = 0mm 0mm 0mm 0mm, width=\textwidth, clip=true]{\fta{}{high}{0}}
+%(b)
+%\end{minipage}
+\caption{Example trajectory snapshots for motions that survey participants rated (a) highly and (b) poorly.
+In (b) the hand ducks under the table in between the first and middle snapshot.}\label{fig:example_trajectory}
+\end{figure}
+
 %\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths

### 2014-10-02 paper2015asml.tex
adding NSF grant #
changing figure sizes
minor mods

@@
 \author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{1}$% <-this % stops a space
-\thanks{*This work was supported by ...}% <-this % stops a space
+\thanks{*This work was partly supported by NSF NRI Grant IIS 1317849}% <-this % stops a space
 \thanks{$^{1}$SRI International, Menlo Park, CA
         {\tt\small arjun.menon@sri.com, sachin.chitta@sri.com}}%
-\thanks{$^{2}$Independent Consultant,
+\thanks{$^{2}$Independent Consultant, Menlo Park, CA
         {\tt\small }}%
@@
 
-{\bf Dynamic Time Warping:}
-
-Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}.
+{\bf Dynamic Time Warping:} Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}.
 It provides a way to compute a distance despite differences in timing of the trajectories as well.
@@
 \begin{figure}[t]
-  \centering{\includegraphics[width=0.99\linewidth]{pictures/meandecrease}}
+  \centering{\includegraphics[width=0.90\linewidth]{pictures/meandecrease}}
   \caption{Random Forest: Mean decrease in accuracy for each feature.}
@@
 \begin{figure}[t]
- \centering{\includegraphics[width=0.99\linewidth]{pictures/gini}}
+ \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
   \caption{Random Forest: Gini index (higher values indicate more important features).}
@@
 \begin{figure}[t]
- \centering{\includegraphics[width=0.99\linewidth]{pictures/gini}}
+ \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
   \caption{Boosting: Gini index (higher values indicate more important features).}

### 2014-10-02 paper2015asml.tex
repositioned the example traj feats

--- 2014-10-02 paper2015asml.tex
@@
 
+\begin{figure}[h]
+\centering
+\begin{tabular}{c|p{0.4\columnwidth}p{0.4\columnwidth}}
+& Low feature-value & High feature-value \\
+\hline
+\rot{wrist\_tool\_dtw} & 
+\adjincludegraphics[trim={.1\width} {.080\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_tool_dtw}{low}{0}} & 
+\adjincludegraphics[trim={.2\width} {.0\height} {.2\width} {.15\height}, clip, width=0.4\columnwidth]{\fta{wrist_tool_dtw}{high}{0}} \\
+
+\rot{tool\_length} & 
+\adjincludegraphics[trim={.1\width} {.2\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{tool_length}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.015\height}, clip, width=0.4\columnwidth]{\fta{tool_length}{high}{0}} \\
+
+\rot{tool\_quat\_dist} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.25\width} {.1\height}, clip, width=0.4\columnwidth]{\fta{tool_quat_dist}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.1\height} {.25\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{tool_quat_dist}{high}{0}} \\
+
+\rot{wrist\_bfs\_dtw} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_bfs_dtw}{low}{0}} & 
+\adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_bfs_dtw}{high}{0}} \\
+\end{tabular}
+\caption{Trajectory examples of the most useful features, where the feature is displayed row-wise and columns are whether the feature is {\em low} or {\em high} for the example.}\label{fig:example_features}
+\end{figure}
+
 \section{Classification}

### 2014-10-02 paper2015asml.tex
conclusions section

@@
 We extended the MoveIt!
-benchmarking architecture~\cite{cohen2012generic} with video recording capabilities to quickly generate a large number of trajectories.
+benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories.
 The MoveIt!
@@
 Both planners are implemented in the OMPL library~\cite{OMPL:2012}.
+All the paths generated by RRT-Connect are shortcutted for a finite amount of time to improve the quality of the paths.
 The planned trajectories are post-processed by an iterative-time parameterization method that time-parameterizes the trajectories while taking into account velocity and accelerations.
@@
 A total of 103 different trajectories were generated in this manner.
+
+Our main focus in generating trajectories was to generate a wide sample.
+The choice of planners used to generate the trajectory was not important.
+We preferred the randomized planners since they natively generate a wide variety of plans that are less likely to be similar to each other.
+We could have chosen to generate all our trajectories with an  asymptotically optimal planner like RRT* but we would not have been able to generate a variety of plans.
+This would also have restricted the richness of the trajectory features we could work with.
+Our intent in this work is not to evaluate particular planning algorithms but instead to examine the preference that human-observers have for generated trajectories regardless of the methods used to generate them.
 
@@
 
+The classification results clearly show the value in using a data-driven approach to determining human preferences for robot trajectories. It is difficult, 
+in obstacle filled cluttered environments, to hand-design an appropriate cost function for generating human preferred trajectories. Tuning coefficients 
+and weights for all the individual cost components for designing preferred trajectories requires significant effort. 
+The classifiers that we have built give us a notion of the quality of trajectories 
+being generated without an explicit need to model the cost function. The output from the classifier could be integrated into gradient-free motion planning schemes 
+(e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
+should also help in the design of cost functions, when necessary, especially since some the features may not be traditionally included as part of a cost function. 
+In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which mostly relies on a path length optimization criterion.
+
+\section{Conclusion}
+In this work, we have generated ratings for a large dataset of robot trajectories. We have already gained significant insights into the 
+types of trajectories that are preferred by human observers. In particular, we have been able to identify a set of features that seem to influence 
+the human's perception of the trajectories the most. This can help us in defining better cost functions when generating the trajectories. The classifiers 
+that we have learned could also be used to directly inform gradient free motion planning algorithms by assigning a rating to new trajectories. Our 
+results were generated using a single robot model (the PR2) but we believe that they will easily translation to other robot models as well. To this end, 
+we will keep the survey open and attempt to collect data for a larger set of trajectories in 
+more environments and with different robots.
+
+The dataset we have collected and the analysis we have already done in this paper will serve as a
+ platform for further development of the idea of using 
+a data-driven approach to motion planning. We intend to attempt to learn a cost function for describing the relation between the features and 
+the ratings of the generated trajectories using inverse reinforcement learning. We hope to examine whether a cost function derived in this manner will 
+lead to generation of more preferred trajectories. We would also like to extend our set of features to include more dynamic features to capture the effect 
+that the dynamics of the motions have on the perception of the motion. The infrastructure that we have developed for generating and conducting 
+surveys is fairly generic and could easily be extended to surveying people using trajectories from real robots. This is another avenue we intend to explore 
+in the near future. We also intend to explore the potential for open-sourcing the data and infrastructure to allow other researchers to conduct surveys of their own.
+
 %\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths

### 2014-10-02 paper2015asml.tex
stylistic changes

--- 2014-10-02 paper2015asml.tex
@@
 \centering
-\begin{minipage}[][][s]{0.45\columnwidth}
-\centering
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}}
-%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{2}}
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}}
-%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{4}}
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}}
-(a)
-\end{minipage}
-\hfill
-\begin{minipage}[][][s]{0.45\columnwidth}
-\centering
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}}
-%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{2}}
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}}
-%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{4}}
-\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}}
-(b)
-\end{minipage}
+\begin{tabular}{c|c}
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}} &
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}} \\
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}} &
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}} \\
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}} &
+\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}} \\
+(a) & (b) \\
+\end{tabular}
+%\begin{minipage}[][][s]{0.45\columnwidth}
+%\centering
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}}
+%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{2}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}}
+%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{4}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}}
+%(a)
+%\end{minipage}
+%\hfill
+%\vline
+%\hfill
+%\begin{minipage}[][][s]{0.45\columnwidth}
+%\centering
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}}
+%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{2}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}}
+%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{4}}
+%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}}
+%(b)
+%\end{minipage}
 \caption{Example trajectory snapshots for motions that survey participants rated (a) highly and (b) poorly.
@@
 \centering
-\begin{tabular}{p{0.3\columnwidth}|p{0.2\columnwidth}|p{0.5\columnwidth}}
+\begin{tabular}{|p{0.3\columnwidth}|p{0.2\columnwidth}|p{0.5\columnwidth}|}
 \hline
@@
 \centering
-\begin{tabular}{c|p{0.4\columnwidth}p{0.4\columnwidth}}
+\begin{tabular}{|c|cc|}%p{0.4\columnwidth}p{0.4\columnwidth}}
+\hline
 & Low feature-value & High feature-value \\
@@
 \adjincludegraphics[trim={.1\width} {.0\height} {.1\width} {.0\height}, clip, width=0.4\columnwidth]{\fta{wrist_bfs_dtw}{high}{0}} \\
+\hline
 \end{tabular}

### 2014-10-02 paper2015asml.tex
fixed quotation marks, removed a few unnecessary spaces

@@
 It is not clear whether minimum-jerk trajectories (or other cost functions) are the best descriptors of motions in such environments or if there are other features that better explain human-observer preferences in such sitations.
-Our goal in this work is to collect the data needed to answer the query, "Which robot trajectories in cluttered environments do human-observers prefer".
-A second question that we would like to answer is "Why are certain robot trajectories preferred by human-observers?".
-This is a larger question that we aim to address in future work and instead we focus on,"Can we predict a human-observer preference for a given trajectory".
-The answer to this latter question, we hope, will eventually point us towards trajectories that are better accepted by human observers.
+Our goal in this work is to collect the data needed to answer the query, ``Which robot trajectories in cluttered environments do human-observers prefer'' A second question that we would like to answer is ``Why are certain robot trajectories preferred by human-observers?'' This is a larger question that we aim to address in future work and instead we focus on, ``Can we predict a human-observer preference for a given trajectory'' The answer to this latter question, we hope, will eventually point us towards trajectories that are better accepted by human observers.
 
@@
 These make up our unary features computed only on the input trajectory.
-We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the "expected" path.
+We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the ``expected'' path.
 We go over the multi-point features in the next section.
@@
 We compute path length for the joint-space trajectory and for three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links of the right arm (the moving arm).
-We also compute the length of the trajectory in angular distance by summing over the waypoints $\arccos( \langle Q_\text{i},Q_\text{i+1}\rangle^2-1 )$ where $Q_\text{i}$ is the quaternion orientation for the $i$-th waypoint (referred to later as the "quat\_dist" feature).
+We also compute the length of the trajectory in angular distance by summing over the waypoints $\arccos( \langle Q_\text{i},Q_\text{i+1}\rangle^2-1 )$ where $Q_\text{i}$ is the quaternion orientation for the $i$-th waypoint (referred to later as the ``quat\_dist'' feature).
 

### 2014-10-02 paper2015asml.tex
more fixes

--- 2014-10-02 paper2015asml.tex
@@
 }
-\caption{The rating categories and the associated scale presented to the survey respondent}
+\caption{The four rating categories/attributes and the associated scales for each presented to the survey respondent.}
 \label{tab:rating_scales}
@@
 The survey was built on Amazon's Mechanical Turk~\cite{paolacci2010running}, which has become a useful tool for large online data gathering and crowd-sourced work.
-Each respondent in the survey was presented with a set of demographic questions, a video, and four scales to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
+Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
+The ratings were later converted into numerical values on a scale of 1 to 7.
 Each respondent was expected to rate multiple trajectories but could not rate a trajectory more than once.
@@
 Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$).
-Confidence intervals for the overall rating of each trajectory are shown in Fig.~\ref{confidence}.
-The small spread in the confidence intervals clearly indicates that different users had similar opinions about the same trajectory.
-
+The 95\% confidence intervals for all the four ratings for each trajectory was less than 1.0.
 \begin{table}
@@
 
+We identified and computed multiple types of trajectory features.
 The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory.
@@
 
-\subsection{Multi-Trajectory Features}
+\subsection{Comparative Trajectory Features}
 \label{subsec:multi_traj_feat}
@@
 }
-\caption{Decoding of feature names into descriptions}
+\caption{The complete list of computed trajectory features used for classification.}
 \end{table}
@@
 \label{fig:feature_importance_1}
+\vspace{-0.2in}
 \end{figure}
-
 \begin{figure}[t]
@@
 \label{fig:feature_importance_2}
+\vspace{-0.3in} 
 \end{figure}
-
 \begin{figure}[t]
@@
 \label{fig:gini_boosting}
+\vspace{-0.3in}
 \end{figure}
@@
 \section{Conclusion}
+\label{sec:conclusion}
 In this work, we have generated ratings for a large dataset of robot trajectories. We have already gained significant insights into the 

### 2014-10-02 paper2015asml.tex

--- 2014-10-02 paper2015asml.tex
@@
 We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the ``expected'' path.
-We go over the multi-point features in the next section.
+We go over these comparative features in the next section.
 First, we detail the features we computed on each trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$.

### 2014-10-02 paper2015asml.tex
delete 3 refs, add break

@@
 Our survey followed standard guidelines for privacy of respondent information when conducting a survey on Mechanical Turk.} where we asked human observers to rate a variety of trajectories of the PR2 robot.
-The survey was built on Amazon's Mechanical Turk~\cite{paolacci2010running}, which has become a useful tool for large online data gathering and crowd-sourced work.
+The survey was built on Amazon's Mechanical Turk, which has become a useful tool for large online data gathering and crowd-sourced work.
 Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
@@
 Respondents on Mechanical Turk are only paid if the survey organizer approves their response.
-A control question, asking the respondent to identify the known action of a robot in a video, served as the first satisficing trap to ensure that respondents were truthfully attempting the survey (see~\cite{krosnick1991response}~\cite{krosnick1996satisficing} for more information on satisficing in survey responses).
+A control question, asking the respondent to identify the known action of a robot in a video, served as the first satisficing trap to ensure that respondents were truthfully attempting the survey (see~\cite{krosnick1996satisficing} for more information on satisficing in survey responses).
 Survey responses were also manually examined to assess whether respondents were responding honestly.
@@
 
-We also use the Interpolation-based Modified Hausdorff distance~\cite{dubuisson1994modified}~\cite{chen2013dynamic}, which is defined as:
+We also use the Interpolation-based Modified Hausdorff distance~\cite{dubuisson1994modified}, which is defined as:
 
@@
 
-{\bf Dynamic Time Warping:} Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}\cite{chen2013dynamic}.
+{\bf Dynamic Time Warping:} Dynamic Time Warping (DTW) was previously used in a wide varity of applications ranging from speech recognition, gesture recognition, handwriting matching, and can be applied here to compare two different trajectories~\cite{senin2008dynamic}.
 It provides a way to compute a distance despite differences in timing of the trajectories as well.
@@
 \end{equation} where $n_A = |\tj{A}|$ and $n_B = |\tj{B}|$ which are the number of waypoints in the trajectory.
-More on the construction of the \tj{DTW} can be read in~\cite{senin2008dynamic}\cite{chen2013dynamic}.
+More on the construction of the \tj{DTW} can be read in~\cite{senin2008dynamic}.
 The DTW feature is the cost of the warping path.
@@
 the tool path) denoted by wrist\_tool\_dtw, the Hausdorf distance computed for the end-effector and the max curvature computed for the elbow.
+
 Using 10-fold cross-validation, we also design a simpler tree that has just three terminal nodes with the wrist\_tool\_dtw feature serving as the key discriminant among the classes.

### 2014-10-02 paper2015asml.tex
reworded feature description table

@@
 \hline
-h\_[link]\_bfs, h\_[link1]\_[link2], H\_[link]\_bfs, H\_[link1]\_[link2] & Cartesian & Assymetric ($h$) or symmetric ($H$) hausdorff distance between 3D cartesian path traced by [link] in \tj{} w.r.t BFS path or between 3D Cartesian path traced by [link1] and [link2] in \tj{}\\
+h\_[link]\_bfs, h\_[link1]\_[link2], H\_[link]\_bfs, H\_[link1]\_[link2] & Cartesian & Assymetric ($h$) or symmetric ($H$) hausdorff distance between 3D path of [link] w.r.t BFS path or between 3D paths of [link1] and [link2]\\
 \hline
-[link]\_bfs\_dtw, [link1]\_[link2]\_dtw & Cartesian & DTW cost between path traced by [link] in \tj{} and the BFS path, or between 3D-cartesian-path traced by [link1] and [link2] in \tj{}\\
+[link]\_bfs\_dtw, [link1]\_[link2]\_dtw & Cartesian & DTW cost between path of [link] and the BFS path, or between paths of [link1] and [link2]\\
 \hline

### 2014-10-02 paper2015asml.tex
minor modes
minor mods

@@
 Decision trees for classification provide a good visual representation especially in the presence of complex features, where the relationships between predictors and response is not immediately obvious.
-The features of interest in this tree include the DTW distance computed for the wrist (vs.
+The features of interest in the first decision tree that we derived include the DTW distance computed for the wrist (vs.
 the tool path) denoted by wrist\_tool\_dtw, the Hausdorf distance computed for the end-effector and the max curvature computed for the elbow.
-
 Using 10-fold cross-validation, we also design a simpler tree that has just three terminal nodes with the wrist\_tool\_dtw feature serving as the key discriminant among the classes.
@@
 Large values of the wrist\_tool\_dtw feature correspond to extra rotations of the wrist as the tool is moving towards the goal - such trajectories are rated as low quality by the human observers.
+
 The second classification framework we use is a random forest ensemble of trees.
@@
 \label{sec:conclusion}
-In this work, we have generated ratings for a large dataset of robot trajectories. We have already gained significant insights into the 
+In this work, we have gathered ratings for a large dataset of robot trajectories. We have already gained significant insights into the 
 types of trajectories that are preferred by human observers. In particular, we have been able to identify a set of features that seem to influence 

### 2014-10-02 paper2015asml.tex
line break
more mods

@@
 \end{tabular}
-\caption{Trajectory examples of the most useful features, where the feature is displayed row-wise and columns are whether the feature is {\em low} or {\em high} for the example.}\label{fig:example_features}
+\caption{Example trajectories for different values of the most useful features.
+Trajectories on the left have low feature values while trajectories on the right have high features values for each corresponding feature in that row.}\label{fig:example_features}
 \end{figure}
@@
 The second measure is based on the Gini index and measures the total decrease in node heterogeneity/impurities from splitting on the feature.
+
 For both measures higher values denote that the features are more important.

### 2014-10-02 paper2015asml.tex
more mods

@@
 In Section~\ref{sec:survey}, we present details of the online survey including information about the types of trajectories and environments used.
-Section~\ref{sec:raw_results} presents the results from the survey and analysis, including the trajectory features used for classification and the classification approaches used with the data.
+Section~\ref{sec:raw_results} presents the results from the survey and analysis, including the trajectory features used for classification and Section~\ref{sec:classification} covers the classification approaches used with the data.
 In Section~\ref{sec:conclusion}, we conclude with a discussion of how our results could be used to design motion planners that generate more human-acceptable trajectories.
@@
   \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
-  \caption{Histogram of averaged ratings for the 103 different trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Inelegant, Smooth vs. Rough, Overall Rating}
+  \caption{Histogram of averaged ratings for the 4 different attributes for 103 trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Awkward, Smooth vs. Rough, Overall Rating}
 \label{fig:survey_raw}
@@
 \section{Classification}
+\label{sec:classification}
 The survey data provides a great platform to answer the question, ``Can we predict a human-observer preference for a given trajectory?''.

### 2014-10-02 paper2015asml.tex
typo
modified the spacing of some figs and tables, removed some lines
spacing and formatting changes

@@
 \scalebox{0.9}{
-\begin{tabular}{|c||c|c|c|c|c|c|c|}
+\begin{tabular}{cccccccc}
+&&&&&&&\\
 \hline
@@
 \label{tab:correlation}
+\vspace{-0.1in}
 \end{table}
@@
 {\bf Hausdorff Distance:}
-The Hausdorff distance $H_\text{AB}$ between trajectory $\xi_\text{A}$ and trajectory $\xi_\text{B}$ can be computed as follows:
+The Hausdorff distance $H_\text{AB}$ between trajectory $\xi_\text{A}$ and $\xi_\text{B}$ can be computed as follows:
 \begin{equation}
@@
 \centering
-\begin{tabular}{|p{0.3\columnwidth}|p{0.2\columnwidth}|p{0.5\columnwidth}|}
+\begin{tabular}{p{0.3\columnwidth}p{0.2\columnwidth}p{0.5\columnwidth}}
+&&\\
 \hline
@@
 \caption{The complete list of computed trajectory features used for classification.}
+\vspace{-0.2in} 
 \end{table}
@@
 \centering
-\begin{tabular}{|c|cc|}%p{0.4\columnwidth}p{0.4\columnwidth}}
+\begin{tabular}{ccc}%p{0.4\columnwidth}p{0.4\columnwidth}}
+&&\\
 \hline
@@
 Trajectories on the left have low feature values while trajectories on the right have high features values for each corresponding feature in that row.}\label{fig:example_features}
+\vspace{-0.2in} 
 \end{figure}
@@
 that we have learned could also be used to directly inform gradient free motion planning algorithms by assigning a rating to new trajectories. Our 
-results were generated using a single robot model (the PR2) but we believe that they will easily translation to other robot models as well. To this end, 
+results were generated using a single robot model (the PR2) but we believe that they will easily translate to other robot models as well. To this end, 
 we will keep the survey open and attempt to collect data for a larger set of trajectories in 

### 2014-10-02 paper2015asml.tex
spacing and formatting changes
added spearman comment

--- 2014-10-02 paper2015asml.tex
@@
 
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
+%\overrideIEEEmargins                                      % Needed to meet printer requirements.
 
@@
 Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$).
+The Spearman rank correlation coefficient is also high and statistically significant.
 The 95\% confidence intervals for all the four ratings for each trajectory was less than 1.0.

### 2015-02-24 paper2015asml.tex
latest changes

--- 2014-10-02 paper2015asml.tex
@@
 In (b) the hand of the robot ducks under the table in between the first and middle snapshot.}\label{fig:example_trajectory}
+\vspace{-0.2in}
 \end{figure}
@@
 \label{tab:rating_scales}
+\vspace{-0.2in}
 \end{table*}
@@
 Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$).
-The 95\% confidence intervals for all the four ratings for each trajectory was less than 1.0.
+The Spearman rank correlation coefficients are also high and statistically significant.
 \begin{table}
@@
 (e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
-should also help in the design of cost functions, when necessary, especially since some the features may not be traditionally included as part of a cost function. 
-In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which mostly relies on a path length optimization criterion.
+should also help in the design of cost functions, especially since some the features may not be traditionally included as part of a cost function. 
+In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which rely on a path length criterion.
 
@@
 that we have learned could also be used to directly inform gradient free motion planning algorithms by assigning a rating to new trajectories. Our 
-results were generated using a single robot model (the PR2) but we believe that they will easily translate to other robot models as well. To this end, 
+results were generated using the PR2 but we believe that they will easily translate to other robot models as well. To this end, 
 we will keep the survey open and attempt to collect data for a larger set of trajectories in 

### 2015-02-26 paper2015asml.tex
made some changes

@@
 \usepackage{wrapfig}
-\usepackage{graphicx,color}
+\usepackage{graphicx,color,soul}
 \usepackage{adjustbox}
@@
 %EDITING COMMANDS
-\newcommand{\hilight}[1]{\colorbox{red}{#1}}
+\newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
 \newcommand{\todo}{\colorbox{red}{TODO}}
@@
 These concepts have been well-studies recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
-In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers.
+In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a cost-map was developed in conjunction with sampling based planning to produce trajectories in cluttered ienvironements.}
 
@@
 We identified and computed multiple types of trajectory features.
+This set of features were selected to characterize the trajectories that were produced for the purpose of identifying the most informative sub-set.
+Developing a larger variety of more complicated and descriptive features is left as an avenue for future work.
+
 The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory.
@@
 We also wanted to design features that captured the difference between the planned trajectory and an expected path for the end-effector in the environment.
-The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D voxel approximation of the planning problem.
+
+{\bf Nominal Trajectory Generation}: \hl{The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D grid approximation of the planning problem.
+We plan the path of the end-effector as if it were a }\emph{free-standing cube} \hl{that is not contrained by the kinematics of the arm to which it is attached.
 The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty.
 Any voxel that is partially occupied is considered to be occupied.
-A 26-connected grid was used in the planning process to find a path in the voxel grid for a cubical shape moving from the start position of the end-effector of the arm to the desired goal position.
-Essentially, this would be the optimal path followed by the end-effector from start to goal if it were modeled as a free-standing cube.
+
+The breadth-first search was conducted using this 26-connected grid to find a path in the voxel grid from the start position of the end-effector of the arm (the }\emph{xyz} \hl{coordinates from forward kinematics) to the desired goal position.
+By ignoring orientation and using BFS, we get a piece-wise collection of straight-line paths in the }{\bf workspace} \hl{from that start to goal points.}
+
 Next, we compute comparison features between $\xi_{\text{BFS}}$ and three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links.
 The comparison features computed between each pair of trajectories are the Hausdorff distances~\cite{dubuisson1994modified} and Dynamic Time Warping~\cite{senin2008dynamic} distances.
+\hl{These features are used because of their applicability for comparing trajectories from location based services}~\cite{chen2013dynamic}.
 
@@
 lead to generation of more preferred trajectories. We would also like to extend our set of features to include more dynamic features to capture the effect 
-that the dynamics of the motions have on the perception of the motion. The infrastructure that we have developed for generating and conducting 
+that the dynamics of the motions have on the perception of the motion.
+\hl{Moreover, the feature set we describe in the paper should grow to further describe the spatial motions even more.} The infrastructure that we have developed for generating and conducting
 surveys is fairly generic and could easily be extended to surveying people using trajectories from real robots. This is another avenue we intend to explore 
@@
 \section*{ACKNOWLEDGMENT}
-We would like to thank Leila Takayama for valuable advice on structuring the online survey. We would like to thank Lydia Kavraki and Mark Moll for valuable discussions and suggestions. 
+We would like to thank Leila Takayama for valuable advice on structuring the online survey.
+We would like to thank Lydia Kavraki,  Mark Moll, and Mike Phillips for valuable discussions and suggestions.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-02-26 paper2015asml.tex
fixed typos

@@
 Predictability implies that the robot will always be consistent in its actions, removing the element of surprise for human observers.
-These concepts have been well-studies recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
+These concepts have been well-studied recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
 In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a cost-map was developed in conjunction with sampling based planning to produce trajectories in cluttered ienvironements.}
@@
 It is not clear whether minimum-jerk trajectories (or other cost functions) are the best descriptors of motions in such environments or if there are other features that better explain human-observer preferences in such sitations.
-Our goal in this work is to collect the data needed to answer the query, ``Which robot trajectories in cluttered environments do human-observers prefer'' A second question that we would like to answer is ``Why are certain robot trajectories preferred by human-observers?'' This is a larger question that we aim to address in future work and instead we focus on, ``Can we predict a human-observer preference for a given trajectory'' The answer to this latter question, we hope, will eventually point us towards trajectories that are better accepted by human observers.
+Our goal in this work is to collect the data needed to answer the query, ``Which robot trajectories in cluttered environments do human-observers prefer'' A second question that we would like to answer is ``Why are certain robot trajectories preferred by human-observers?'' This is a larger question that we aim to address in future \hl{work.
+Instead we} focus on answering ``Can we predict a human-observer preference for a given trajectory'' The answer to this latter question, we hope, will eventually point us towards trajectories that are better accepted by human observers.
 
@@
 The survey was built on Amazon's Mechanical Turk, which has become a useful tool for large online data gathering and crowd-sourced work.
-Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
+Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficiency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
 The ratings were later converted into numerical values on a scale of 1 to 7.
@@
 \caption{Example trajectories for different values of the most useful features.
-Trajectories on the left have low feature values while trajectories on the right have high features values for each corresponding feature in that row.}\label{fig:example_features}
+\hl{Trajectories on the left have low-end value while trajectories on the right have high-end value for each corresponding feature in that row.}}\label{fig:example_features}
 \vspace{-0.2in} 
@@
 This involves randomly dividing the data into several bootstrap training samples and building a tree on each bootstrap sample.
-Within each boostrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split.
+Within each bootstrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split.
 This brings additional randomness within each tree.

### 2015-02-27 paper2015asml.tex

@@
 These concepts have been well-studied recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
-In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a cost-map was developed in conjunction with sampling based planning to produce trajectories in cluttered ienvironements.}
+In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a cost-map was developed in conjunction with sampling based planning to produce trajectories in cluttered environments.}
 
@@
 A total of 103 different trajectories were generated in this manner.
+Note that although the planners we use are randomized, the post-processing step (short-cutting) helps significantly in smoothing the plans.
 
@@
 Our survey followed standard guidelines for privacy of respondent information when conducting a survey on Mechanical Turk.} where we asked human observers to rate a variety of trajectories of the PR2 robot.
-The survey was built on Amazon's Mechanical Turk, which has become a useful tool for large online data gathering and crowd-sourced work.
+\hl{Our intent with the survey was to get measures of how comfortable the observers were with the generated trajectories.} The survey was built on Amazon's Mechanical Turk, which has become a useful tool for large online data gathering and crowd-sourced work.
 Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficiency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
@@
 This set of features were selected to characterize the trajectories that were produced for the purpose of identifying the most informative sub-set.
-Developing a larger variety of more complicated and descriptive features is left as an avenue for future work.
+\hl{The selected features measure geometric, dynamic, environment and comparative properties of the trajectories.
+Since the observed robot was simulated, these visual features seemed to offer the best measures of feedback that the observers might be using in their rating.} Developing a larger variety of more complicated and descriptive features is left as an avenue for future work.
 
@@
 The maximin distance to joint limits is the maximum over waypoints, of the miniminum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$.
+Note that although the joint limit feature may not be immediately perceived by users, we chose to include it to see if the data might prove otherwise.
 
@@
 
-In all, we utilized three different classification frameworks to predict overall ratings - decision trees, random forests and boosting.
+\hl{The goal of the classification problem is to predict the overall rating that human observers would assign to a new trajectory based on the observed ratings in the survey.} In all, we utilized three different classification frameworks to predict overall ratings - decision trees, random forests and boosting.
 As a base framework we first construct a decision tree model.
@@
 the ratings of the generated trajectories using inverse reinforcement learning. We hope to examine whether a cost function derived in this manner will 
-lead to generation of more preferred trajectories. We would also like to extend our set of features to include more dynamic features to capture the effect 
+lead to generation of more preferred trajectories.
+
+\hl{We would also like to extend our set of features in multiple ways. We would like to include more dynamic features to capture the effect 
 that the dynamics of the motions have on the perception of the motion.
-\hl{Moreover, the feature set we describe in the paper should grow to further describe the spatial motions even more.} The infrastructure that we have developed for generating and conducting
-surveys is fairly generic and could easily be extended to surveying people using trajectories from real robots. This is another avenue we intend to explore 
+We would also like to introduce more environment specific features, e.g.
+the descriptors in~\cite{10-jetchev-ICRA}, to account better for the effect of the environment on the ratings.
+Context and task specific features would be useful in capturing task specific motions, e.g.
+a robot turning a screwdriver may actually need to turn its wrist continuously.
+Incorporating task motions and relevant environments, e.g.
+picking up an object from a cluttered table, would also help in evaluating features like legibility.
+We would also like to examine how people's ratings change as they spend more time observing the robot (this effort will require a longer running survey or multiple surveys spaced apart in time).
+We would also like to include features to encode the choice of start and goal positions for the robot - several of the existing features could easily be adapated to take this into account.} The infrastructure that we have developed for generating and conducting
+surveys is fairly generic and could also easily be extended to surveying people using trajectories from real robots. This is another avenue we intend to explore 
 in the near future. We also intend to explore the potential for open-sourcing the data and infrastructure to allow other researchers to conduct surveys of their own.

### 2015-02-27 paper2015asml.tex
fix for highlight
additions, notes

@@
 These concepts have been well-studied recently with the intent of measuring legibility~\cite{Lichtenthaler:2011}, using expressions to improve the readability of robots~\cite{Takayama:2011} and generating anticipation~\cite{Gielniak:2011}.
-In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a cost-map was developed in conjunction with sampling based planning to produce trajectories in cluttered environments.}
+In~\cite{Dragan:2013}, the notion of legibility was incorporated into a constrained trajectory optimization motion planner to generate motions that are more legible to human observers,\hl{ while in}~\cite{mainprice2011planning} \hl{a costmap was developed for use in motion planning taking into account human interaction constraints like gaze, proximity to the robot and ease of hand-off.}
 
@@
 We would also like to introduce more environment specific features, e.g.
-the descriptors in~\cite{10-jetchev-ICRA}, to account better for the effect of the environment on the ratings.
+the descriptors in}~\cite{10-jetchev-ICRA}, \hl{to account better for the effect of the environment on the ratings.
 Context and task specific features would be useful in capturing task specific motions, e.g.

### 2015-02-27 paper2015asml.tex
close to final version

@@
 \newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
+\renewcommand{\hl}[1]{{#1}}
 \newcommand{\todo}{\colorbox{red}{TODO}}
@@
 
-We extended the MoveIt!
-benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories.
+We extended the MoveIt!~\cite{MoveIt} benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories.
 The MoveIt!
@@
 
+We computed the correlation between observer ratings of smoothness vs.
+the measured smoothness of the trajectories (Table~\ref{tab:corr_smoothness}).
+Note that higher values of the measured smoothness feature  indicate that the trajectories are jagged or non-smooth.
+The correlation between the smoothness rating and the measured smoothness is not very high, indicating that human observer perception of smoothness is different from the particular measure that we used for the robot trajectories.
+A likely explanation is that excessive motions of the wrist (while still measured as smooth) are perceived as non-smooth by the human observer.
+We also computed the correlation between path lengths and the efficiency rating assigned by the human observer (Table~\ref{tab:path_length}).
+Here, the correlation was high (the negative value of the correlation coefficients indicate that higher path lengths are perceived as less efficient).
+
+\begin{table}
+\centering
+\begin{tabular}{|p{1.25cm}|c|c|c|c|}
+\hline
+& Class & Smoothness & Smoothness & Smoothness \\ 
+& Smooth & (Joint) & (Tool) & (Wrist) \\ \hline
+Class Smooth & 1 & -0.27 & -0.29 & -0.38 \\ \hline
+Smoothness (Joint) & -0.27 & 1 & 0.68 & 0.62 \\ \hline
+Smoothness (Tool) & -0.29 & 0.68 & 1 & 0.87 \\ \hline
+Smoothness (Wrist) & -0.38 & 0.62 & 0.87 & 1 \\ \hline
+\end{tabular}
+\caption{Correlation between human observed rating for smoothness and measured smoothness values for the trajectories.}
+\label{tab:corr_smoothness}
+\vspace{-0.1in}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|p{1.25cm}|c|c|c|c|}
+\hline
+& Class & Path Length & Path Length & Path Length \\ 
+& Efficient & (Joint) & (Wrist) & (Tool) \\ \hline
+\hline
+Class Efficient & 1 & -0.79 & -0.8 & -0.76 \\ \hline
+Path Length (Joint) & -0.79 & 1 & 0.89 & 0.86 \\ \hline
+Path Length (Wrist) & -0.8 & 0.89 & 1 & 0.97 \\ \hline
+Path Length (Tool) & -0.76 & 0.86 & 0.97 & 1 \\ \hline
+\end{tabular}
+\caption{Correlation between human observed rating for efficiency and measured path lengths for the trajectories.}
+\label{tab:path_length}
+\vspace{-0.1in}
+\end{table}
+
 \subsection{Trajectory Features}
@@
 The comparison features computed between each pair of trajectories are the Hausdorff distances~\cite{dubuisson1994modified} and Dynamic Time Warping~\cite{senin2008dynamic} distances.
-\hl{These features are used because of their applicability for comparing trajectories from location based services}~\cite{chen2013dynamic}.
+\hl{These features have also often been used for comparing trajectories in applications like location based services and are useful in comparing trajectories that might have different lengths or shapes.}~\cite{chen2013dynamic}.
 
@@
 a data-driven approach to motion planning. We intend to attempt to learn a cost function for describing the relation between the features and 
-the ratings of the generated trajectories using inverse reinforcement learning. We hope to examine whether a cost function derived in this manner will 
+the ratings of the generated trajectories using inverse reinforcement learning (e.g.
+as in~\cite{kalakrishnanICRA2013},~\cite{abbeel2004apprenticeship}).
+We hope to examine whether a cost function derived in this manner will
 lead to generation of more preferred trajectories.

### 2015-02-27 paper2015asml.tex
adding note on why legibility is not important

@@
 Each respondent in the survey was presented with a set of demographic questions, a video, and asked to rate the robot's motion on the attributes  of elegance, efficiency, smoothness and overall impression (a summary of the rating scales used in the main survey question can be seen in Table~\ref{tab:rating_scales}).
-The ratings were later converted into numerical values on a scale of 1 to 7.
+\hl{The users were shown the start and desired goal positions of the robot before they viewed the video of the motion.
+Thus, the users had prior information about the robot's intent.} The ratings were later converted into numerical values on a scale of 1 to 7.
 Each respondent was expected to rate multiple trajectories but could not rate a trajectory more than once.
@@
 Incorporating task motions and relevant environments, e.g.
-picking up an object from a cluttered table, would also help in evaluating features like legibility.
+picking up an object from a cluttered table, would also help in evaluating features like legibility (note that legibility does not have a clear meaning in our experiments since the users already knew the robot's intent).
 We would also like to examine how people's ratings change as they spend more time observing the robot (this effort will require a longer running survey or multiple surveys spaced apart in time).

### 2015-02-27 paper2015asml.tex
fixed reference to finger-tip/tool

@@
 %\end{equation}
-We compute path length for the joint-space trajectory and for three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links of the right arm (the moving arm).
+We compute path length for the joint-space trajectory and for three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip/tool ($\xi_\text{tip}$) links of the right arm (the moving arm).
 We also compute the length of the trajectory in angular distance by summing over the waypoints $\arccos( \langle Q_\text{i},Q_\text{i+1}\rangle^2-1 )$ where $Q_\text{i}$ is the quaternion orientation for the $i$-th waypoint (referred to later as the ``quat\_dist'' feature).
@@
 
-Next, we compute comparison features between $\xi_{\text{BFS}}$ and three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip ($\xi_\text{tip}$) links.
+Next, we compute comparison features between $\xi_{\text{BFS}}$ and three workspace trajectories for the elbow ($\xi_\text{elbow}$), wrist ($\xi_\text{wrist}$) and finger-tip/tool ($\xi_\text{tip}$) links.
 The comparison features computed between each pair of trajectories are the Hausdorff distances~\cite{dubuisson1994modified} and Dynamic Time Warping~\cite{senin2008dynamic} distances.

### 2015-02-28 paper2015asml.tex
adding movie

@@
 \usepackage{adjustbox}
+\usepackage{flushend}
 
@@
 \title{\LARGE \bf
-Towards a Data-driven Approach to Motion Planning
+Towards a Data-driven Approach to Human Preferences in Motion Planning
 }
 
-\author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{1}$% <-this % stops a space
+\author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{3}$% <-this % stops a space
 \thanks{*This work was partly supported by NSF NRI Grant IIS 1317849}% <-this % stops a space
-\thanks{$^{1}$SRI International, Menlo Park, CA
-        {\tt\small arjun.menon@sri.com, sachin.chitta@sri.com}}%
+\thanks{$^{1}$Anki Robotics, San Francisco, CA - this work was performed while Arjun was at SRI International
+        {\tt\small }}%
 \thanks{$^{2}$Independent Consultant, Menlo Park, CA
         {\tt\small }}%
+\thanks{$^{3}$Corresponding author: SRI International, Menlo Park, CA
+        {\tt\small sachin.chitta@sri.com}}%
 }
@@
 
-We extended the MoveIt!~\cite{MoveIt} benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories.
-The MoveIt!
-architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries).
+We extended the MoveIt!
+benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories.
+The MoveIt!~\cite{MoveIt}  architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries).
 It also allows for the use of any supported planning algorithms (planners).
@@
 We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen (Countertop), an industrial workbench (Industrial) and a drawer on a table (Drawer).
-Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot (Fig.~\ref{fig:PR2}.
+Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot (Fig.~\ref{fig:PR2}).
 The robot was positioned in the scenes manually and motion planning queries were constructed by hand.
@@
 RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: PathLength and MaxMinClearance.The objective for PathLength minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments).
-The objective for MaxMinClearance computes uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize.
+The objective for MaxMinClearance uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize.
 Both planners are implemented in the OMPL library~\cite{OMPL:2012}.
@@
 It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well.
-An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall).
+An analysis of the ratings for each trajectory showed that the four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall).
 Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes.
@@
 Note that higher values of the measured smoothness feature  indicate that the trajectories are jagged or non-smooth.
-The correlation between the smoothness rating and the measured smoothness is not very high, indicating that human observer perception of smoothness is different from the particular measure that we used for the robot trajectories.
+The smoothness rating is not very highly correlated with any of the measured smoothness features.
 A likely explanation is that excessive motions of the wrist (while still measured as smooth) are perceived as non-smooth by the human observer.
@@
 \hline
-& Class & Smoothness & Smoothness & Smoothness \\ 
-& Smooth & (Joint) & (Tool) & (Wrist) \\ \hline
-Class Smooth & 1 & -0.27 & -0.29 & -0.38 \\ \hline
+& Smooth & Smoothness & Smoothness & Smoothness \\ 
+& & (Joint) & (Tool) & (Wrist) \\ \hline
+Smooth & 1 & -0.27 & -0.29 & -0.38 \\ \hline
 Smoothness (Joint) & -0.27 & 1 & 0.68 & 0.62 \\ \hline
@@
 \centering
-\begin{tabular}{|p{1.25cm}|c|c|c|c|}
+\begin{tabular}{|p{1.5cm}|c|c|c|c|}
 \hline
-& Class & Path Length & Path Length & Path Length \\ 
-& Efficient & (Joint) & (Wrist) & (Tool) \\ \hline
-\hline
-Class Efficient & 1 & -0.79 & -0.8 & -0.76 \\ \hline
+& Efficient & Path Length & Path Length & Path Length \\ 
+&  & (Joint) & (Wrist) & (Tool) \\ \hline
+Efficient & 1 & -0.79 & -0.8 & -0.76 \\ \hline
 Path Length (Joint) & -0.79 & 1 & 0.89 & 0.86 \\ \hline
@@
 These make up our unary features computed only on the input trajectory.
-We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the ``expected'' path.
+We also include features that compare each trajectory to a nominal trajectory - these features serve to characterize how much the trajectory deviates from the ``expected'' path.
 We go over these comparative features in the next section.
@@
 {\bf Joint Limit Distance: }
-The maximin distance to joint limits is the maximum over waypoints, of the miniminum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$.
+The maxmin distance to joint limits is the maximum over waypoints, of the minimum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$.
 Note that although the joint limit feature may not be immediately perceived by users, we chose to include it to see if the data might prove otherwise.
@@
 {\bf Nominal Trajectory Generation}: \hl{The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D grid approximation of the planning problem.
-We plan the path of the end-effector as if it were a }\emph{free-standing cube} \hl{that is not contrained by the kinematics of the arm to which it is attached.
+We plan the path of the end-effector as if it were a }\emph{free-standing cube} \hl{that is not constrained by the kinematics of the arm to which it is attached.
 The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty.
@@
 For every tree constructed from a bootstrapped training sample there is one Out of Bag (OOB) error which can be thought of as a test error for every tree.
-It is estimated internally, during the run.
-On test data, the random forest method returns 83\% accuracy.
+Our model produced an OOB estimate of error rate equivalent to 19\%.
+On the test data, the random forest method returns 83\% accuracy.
 
@@
  \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
-  \caption{Random Forest: Gini index (higher values indicate more important features).}
+  \caption{Random Forest: Gini index based relative importance measure.}
 \label{fig:feature_importance_2}
@@
 \begin{figure}[t]
- \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
-  \caption{Boosting: Gini index (higher values indicate more important features).}
+ \centering{\includegraphics[width=0.90\linewidth]{pictures/gini_boosting}}
+  \caption{Boosting: Gini index based relative importance measure.}
 \label{fig:gini_boosting}
@@
 (e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
-should also help in the design of cost functions, especially since some the features may not be traditionally included as part of a cost function. 
+should also help in the design of cost functions, especially since some of the features may not be traditionally included as part of a cost function. 
 In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which rely on a path length criterion.
@@
 more environments and with different robots.
+
+%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+                                  % on the last page of the document manually. It shortens
+                                  % the textheight of the last page by a suitable amount.
+                                  % This command does not take effect until the next page
+                                  % so it should come on the page before the last. Make
+                                  % sure that you do not shorten the textheight too much.
 
@@
 
-%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
-                                  % on the last page of the document manually. It shortens
-                                  % the textheight of the last page by a suitable amount.
-                                  % This command does not take effect until the next page
-                                  % so it should come on the page before the last. Make
-                                  % sure that you do not shorten the textheight too much.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
